---
title: "MSDS 597 HW7"
author: "Shiqin Cai"
date: "November 3, 2016"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(stringr)
library(choroplethr)
library(choroplethrMaps)
library(ggthemes)
```


First, we downlad the `mortality.csv` file and save it into the folder named `data`.

```{r, include=FALSE}
mortality <- read_csv("data/mortality.csv")
```


We then clean the `mortality` as the folllowing steps.

```{r}
mortality$Year <- mortality$Year %>% str_sub(end = 4) %>% as.integer()
mortality$Mortality.num <- mortality$`Maternal mortality ratio (per 100 000 live births)` %>%
  str_extract("[0-9 ]+\\[") %>% str_replace_all("[ \\[]", "") %>% as.integer()
mortality$Lower_bound <- mortality$`Maternal mortality ratio (per 100 000 live births)` %>%
  str_extract("\\[[ 0-9]+") %>% str_replace_all("[ \\[]", "") %>% as.integer()
mortality$Upper_bound <- mortality$`Maternal mortality ratio (per 100 000 live births)` %>%
  str_extract("-[ 0-9]+") %>% str_replace_all("[ -]", "") %>% as.integer()
mortality <- select(mortality, -`Maternal mortality ratio (per 100 000 live births)`)
head(mortality)
```


<!--- does not work because NA value can not be spit into NA;NA;NA, so error in the map_chr
```{r, eval=FALSE}
mortality$`Maternal mortality ratio (per 100 000 live births)` <- 
  mortality$`Maternal mortality ratio (per 100 000 live births)` %>% 
  str_replace_all("[ \\]]", "") %>% str_replace_all("[\\[\\]-]", ";")
# wouldn't work because NA value can not be spit into NA;NA;NA
mortality$Num <- as.integer(map_chr(str_split(mortality$`Maternal mortality ratio (per 100 000 live births)`, pattern=";"), 1))
mortality$Lower_bound <- as.integer(map_chr(str_split(mortality$`Maternal mortality ratio (per 100 000 live births)`, pattern=";", na.rm=T), 2))
mortality$Upper_bound <- as.integer(map_chr(str_split(mortality$`Maternal mortality ratio (per 100 000 live births)`, pattern=";"), 3)) 
```
--->


Next, we download the `income.csv` file and save it into the `data` folder.

```{r, include=FALSE}
income <- read_csv("data/income.csv", skip = 1)
```


We then clean the `income` data as shown below.

```{r}
income$`2013` <- as.integer(str_replace_all(income$`2013`, " ", ""))
income <- income %>% gather(`2013`:`1990`, key="Year", value="Income")
income$Year <- as.integer(income$Year)
head(income)
```


After cleaning both data sets, we merge them by `Country` and `Year`, and only retain rows that have non-missing per capital income values and either non-missing maternal mortality values or non-missing percentage attended births. This results in the new data frame `df`.

```{r}
df <- merge(mortality, income, by=c("Country", "Year"))
df_cleaned <- df %>% filter(!is.na(Income) & (!is.na(Mortality.num) | !is.na(`Births attended by skilled health personnel (%)`)))
```


Since only year 1990 and 2000 contains non-missing value, the following graph displays the number of maternal mortality against income. And the plots below show that countries with lower gross income have greater number of maternal mortality. Comparing the number of mortality in 2000 with that in 1990, we see that the number of maternal mortality decreased in most countries because their national gross income increased in 2000.

```{r}
# income, maternal mortality
df_cleaned %>% filter(!is.na(Income) & !is.na(Mortality.num)) %>% 
  ggplot(aes(x=Income, y=Mortality.num, fill=Country)) +
  geom_point(aes(size=Income), pch=21, show.legend = F) +
  facet_wrap(~ Year) + theme_bw() +
  ggtitle("Income vs Number of Mortality in 1990 and 2000") +
  ylab("Number of Maternal Mortality")
```


The following graph plots the percentage of births atteded by skilled health personnel against income between 2004 and 2013. From the graph we see that the even some countries with low gross national income have high percentage of births attended by skill health personnel in the 20th century. After 2010, countries with higher gross national income have higher percentage of attended births then countries with lower gorss national income.

```{r}
# income, percentage attended births
df_cleaned %>% filter(!is.na(Income) & !is.na(`Births attended by skilled health personnel (%)`)) %>% 
  ggplot(aes(x=Income, y=`Births attended by skilled health personnel (%)`, fill=Country)) +
  geom_point(pch=21, show.legend = F) + theme_bw() +
  facet_wrap(~ Year) + ggtitle("Income vs Percentage of Attended Births") 
```
 

We save a `clean_df.csv` file with the merged, cleaned, and tidy data in the folder named `data`.

```{r}
write_csv(df_cleaned, path = "data/cleaned_df.csv")
```


We only keep the most recent year record for each country using the `max(row_number())`. Since the first letter of the country names in the data frame are capitalized, we need to convert the country names to lower case. We also need to remove the `-` in the country names, and trim them.

```{r}
df_country_demographics <- df_cleaned %>% group_by(Country) %>% filter(row_number()==max(row_number()))
df_country_demographics$Country <- df_country_demographics$Country %>%
  tolower() %>% str_replace("\\([a-z ]+\\)", "") %>% trim() %>%
  str_replace_all("-", " ")
```


The world map of maternal mortality using the most recent year of each country is shown below. From the map, we see that the number of maternal mortality is highest in Africa, while United States has relatively low number of maternal mortality.

```{r, warning=F}
df_country_demographics %>% dplyr::rename(region = Country, value = Mortality.num) %>%
  country_choropleth(title = "Maternal Mortality", num_colors = 9)
```











