---
title: "MSDS 597 HW7"
author: "Shiqin Cai"
date: "November 3, 2016"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(stringr)
library(choroplethr)
library(choroplethrMaps)
library(ggthemes)
```



```{r, include=FALSE}
mortality <- read_csv("mortality.csv")
```

We clean the `mortality` as the folllowing steps.

```{r}
mortality$Year <- mortality$Year %>% str_sub(end = 4) %>% as.integer()
mortality$Mortality.num <- mortality$`Maternal mortality ratio (per 100 000 live births)` %>%
  str_extract("[0-9 ]+\\[") %>% str_replace_all("[ \\[]", "") %>% as.integer()
mortality$Lower_bound <- mortality$`Maternal mortality ratio (per 100 000 live births)` %>%
  str_extract("\\[[ 0-9]+") %>% str_replace_all("[ \\[]", "") %>% as.integer()
mortality$Upper_bound <- mortality$`Maternal mortality ratio (per 100 000 live births)` %>%
  str_extract("-[ 0-9]+") %>% str_replace_all("[ -]", "") %>% as.integer()
head(mortality)
```


<!--- does not work because NA value can not be spit into NA;NA;NA, so error in the map_chr
```{r, eval=FALSE}
mortality$`Maternal mortality ratio (per 100 000 live births)` <- 
  mortality$`Maternal mortality ratio (per 100 000 live births)` %>% 
  str_replace_all("[ \\]]", "") %>% str_replace_all("[\\[\\]-]", ";")
# wouldn't work because NA value can not be spit into NA;NA;NA
mortality$Num <- as.integer(map_chr(str_split(mortality$`Maternal mortality ratio (per 100 000 live births)`, pattern=";"), 1))
mortality$Lower_bound <- as.integer(map_chr(str_split(mortality$`Maternal mortality ratio (per 100 000 live births)`, pattern=";", na.rm=T), 2))
mortality$Upper_bound <- as.integer(map_chr(str_split(mortality$`Maternal mortality ratio (per 100 000 live births)`, pattern=";"), 3)) 
```
--->


```{r, include=FALSE}
income <- read_csv("income.csv", skip = 1)
```

We then clean the `income` data as shown below.

```{r}
income$`2013` <- as.integer(str_replace_all(income$`2013`, " ", ""))
income <- income %>% gather(`2013`:`1990`, key="Year", value="Income")
income$Year <- as.integer(income$Year)
head(income)
```


After cleaning both data sets, we merge them by `Country` and `Year`, and only retain rows that have non-missing per capital income values and either non-missing maternal mortality values or non-missing percentage attended births. This results in the new data frame `df`.

```{r}
df <- merge(mortality, income, by=c("Country", "Year"))
df <- df %>% filter(!is.na(Income) & (!is.na(Mortality.num) | !is.na(`Births attended by skilled health personnel (%)`)))
```


Since only year 1990 and 2000 contains non-missing value, the following graph displays the number of maternal mortality against income. And the plots below show that  

```{r}
# income, maternal mortality
df %>% filter(!is.na(Income) & !is.na(Mortality.num)) %>% 
  ggplot(aes(x=Income, y=Mortality.num, fill=Country)) +
  geom_point(aes(size=Income), pch=21, show.legend = F) +
  facet_wrap(~ Year) + theme_bw() +
  ggtitle("Income vs Number of Mortality in 1990 and 2000") +
  ylab("Number of Maternal Mortality")
```


```{r}
# income, percentage attended births
df %>% filter(!is.na(Income) & !is.na(`Births attended by skilled health personnel (%)`)) %>% 
  ggplot(aes(x=Income, y=`Births attended by skilled health personnel (%)`, fill=Country)) +
  geom_point(aes(size=Income), pch=21, show.legend = F) +
  facet_wrap(~ Year) + ggtitle("Income vs Percentage of Attended Births")
```



```{r}
write_csv(df, path = "cleaned_df.csv")


```


